---
title: "CORONA"
output: html_document
date: "2023-09-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(shiny)
library(shinyjs)
library(leaflet)
```

# load data
```{r, include=FALSE}
facs <- read_csv("facilities.csv")[-1]
missiles <- read_csv("missiles.csv")[-1]
fac_caps_no_unknown <- read_csv("fac_captures.csv")[-1]
fac_caps_with_unknown <- read_csv("fac_caps_with_unknown.csv")[-1]
miss_caps_orig <- read_csv("miss_captures.csv")[-1]

# get date range
min_date = min(append(facs$start_date, missiles$start_date), na.rm = TRUE)
max_date = max(append(facs$start_date, missiles$start_date), na.rm = TRUE)
```

## Per-Facility Visitation Rate
Generate a table with the counts of visits for each facility; visitation leaderboard

```{r}
fac_caps <- fac_caps_with_unknown %>%
  mutate(`Abbreviated Mission` = substr(Mission,1,4))

capture_counts <- fac_caps %>% 
  group_by(facility_name, `Acquisition Date`, `Abbreviated Mission`, start_date) %>%
  summarise(n_caps = n()) 

totals <- capture_counts %>%
  ungroup() %>%
  group_by(facility_name) %>%
  summarise(n_caps = sum(n_caps))
  
totals %>%
  arrange(-n_caps)
```


Could chunk this by country/region? Or by time period (maybe organized around surveillance epochs?)

Also make the ability to generate a plot with the yearly visitation of each site

```{r}
# make a plot of capture activity of Laboratory No. 2, Moscow

# first, need captures per year
yearly_counts <- capture_counts %>%
  mutate(`Year` = as.numeric(substr(`Acquisition Date`,1,4))) %>%
  group_by(facility_name, Year) %>%
  summarise(n_caps = sum(n_caps)) %>%
  ungroup()

fac <- "Tajoura Enrichment Facility"

yearly_counts %>% 
  filter(facility_name == fac) %>%
  ggplot(aes(x = Year, y = n_caps, group = 1)) +
  geom_col()

yearly_counts %>% 
  filter(facility_name == fac) %>%
  ggplot(aes(x = Year, y = n_caps, group = 1)) +
  geom_line()
```

It looks like by far the most captures occur in the early 60s, especially in 1964:

```{r}
yearly_counts %>% group_by(Year) %>%
  summarise(n_caps = sum(n_caps)) %>%
  ggplot(aes(x = Year, y = n_caps)) +
  geom_col()
```

How about the number of photos total, not just captures?

```{r}
sat <- read_csv("sat.csv")

n_pics_per_year <- sat %>% mutate(`Year` = as.numeric(substr(`Acquisition Date`,1,4))) %>%
  group_by(Year) %>%
  summarise(n_pics = n())

n_pics_per_year %>%
  ggplot(aes(x = Year, y = n_pics)) +
  geom_col()
```
It's smoother, but not totally smooth. Let's break it down by data source.

```{r}
n_pics_per_year_grouped <- sat %>% mutate(`Year` = as.numeric(substr(`Acquisition Date`,1,4))) %>%
  group_by(Year, `Data Source`) %>%
  summarise(n_pics = n())

n_pics_per_year_grouped %>%
  ggplot(aes(x = Year)) +
  # geom_ribbon(aes(ymin = 0, ymax = n_pics, fill = `Data Source`), 
  #             alpha = 0.6) + 
  geom_area(aes(y = n_pics, fill = `Data Source`), 
              alpha = 0.9) + 
  scale_fill_manual(values = c("declass1" = "#7eb0d5", "declass2" = "#fd7f6f", "declass3" = "#01A66F"))
```

```{r}
caps_by_year_source <- fac_caps %>% 
  group_by(facility_name, `Acquisition Date`, `Abbreviated Mission`, start_date, `Data Source`) %>%
  summarise(n_caps = n()) %>%
  mutate(`Year` = as.numeric(substr(`Acquisition Date`,1,4))) %>%
  group_by(Year, `Data Source`) %>%
  summarise(n_caps = sum(n_caps)) %>%
  ungroup()

caps_by_year_source %>%
  ggplot(aes(x = Year)) +
  geom_area(aes(y = n_caps, fill = `Data Source`), 
              alpha = 0.9) + 
  scale_fill_manual(values = c("declass1" = "#7eb0d5", "declass2" = "#fd7f6f", "declass3" = "#01A66F"))
```

# average geometry footprint area by dataset

# average resolution by dataset

Worth noting that there are 14 pics with "year = NA".

```{r}
no_geom <- sat %>% filter(geometry == "POLYGON ((0 0,0 0,0 0,0 0,0 0))")

no_geom %>%
  group_by(`Data Source`) %>%
  summarise(count = n())
```



## Calculate and Plot Facility Coverage
```{r}
fac_caps <- fac_caps_orig

# Build coverage DF
coverage <- data.frame(year = integer(),
                       num_extant_facilities = integer(),
                       num_spotted_facilities = integer(),
                       coverage = double())

spotted <- c()

# loop thru years
for (year in as.integer(format(min_date, "%Y")):as.integer(format(max_date, "%Y"))) {
  year_date <- as.Date(paste(year, "-01-01", sep = ""))
  
  # extant facilities
  fac_exist <- facs %>%
    filter(start_date <= year_date) %>%
    distinct(facility_name)  # Get unique facility names
  
  n_fac_exist <- nrow(fac_exist)
  
  # spotted facilities
  fac_spotted <- fac_caps %>%
    filter(`Acquisition Date` <= year_date) %>%
    distinct(facility_name)  # Get unique facility names
  
  spotted <- unique(append(spotted,fac_spotted$facility_name))
  n_fac_spotted <- length(spotted)
  
  # remove earlier facilities for next loop
  fac_caps <- fac_caps %>% 
    filter(`Acquisition Date` > year_date)
  
  coverage <- bind_rows(coverage, data.frame(year = year,
                                             num_extant_facilities = n_fac_exist,
                                             num_spotted_facilities = n_fac_spotted,
                                             coverage = n_fac_spotted / n_fac_exist))
}

## Plot
ggplot(coverage, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = num_extant_facilities, fill = "Extant Facilities"), alpha = 0.75) +
  geom_ribbon(aes(ymin = 0, ymax = num_spotted_facilities, fill = "Spotted Facilities"), alpha = 0.75) +
  scale_fill_manual(values = c("Extant Facilities" = "#7eb0d5", "Spotted Facilities" = "#fd7f6f")) +
  labs(y = "Coverage", 
       x = "Year",
       fill = "") +
  theme_minimal()
```

## Calculate and Plot Missile Coverage
```{r}
miss_caps <- miss_caps_orig

# Build coverage DF
coverage <- data.frame(year = integer(),
                       num_extant_facilities = integer(),
                       num_spotted_facilities = integer(),
                       coverage = double())

spotted <- c()

for (year in as.integer(format(min_date, "%Y")):as.integer(format(max_date, "%Y"))) {
  year_date <- as.Date(paste(year, "-01-01", sep = ""))
  
  # extant facilities
  miss_exist <- missiles %>%
    filter(start_date <= year_date) %>%
    distinct(address_found)  # Get unique facility names
  
  n_miss_exist <- nrow(miss_exist)
  
  # spotted missiles
  miss_spotted <- miss_caps %>%
    filter(`Acquisition Date` <= year_date) %>%
    distinct(address_found)  # Get unique missile names
  
  spotted <- unique(append(spotted,miss_spotted$address_found))
  
  miss_caps <- miss_caps %>% 
    filter(`Acquisition Date` > year_date)
  
  n_miss_spotted <- length(spotted)
  
  coverage <- bind_rows(coverage, data.frame(year = year,
                                             num_extant_missiles = n_miss_exist,
                                             num_spotted_missiles = n_miss_spotted,
                                             coverage = n_miss_spotted / n_miss_exist))
}

## Plot
ggplot(coverage, aes(x = year)) +
  geom_ribbon(aes(ymin = 0, ymax = num_extant_missiles, fill = "Extant missiles"), alpha = 0.75) +
  geom_ribbon(aes(ymin = 0, ymax = num_spotted_missiles, fill = "Spotted missiles"), alpha = 0.75) +
  scale_fill_manual(values = c("Extant missiles" = "#7eb0d5", "Spotted missiles" = "#fd7f6f")) +
  labs(y = "Coverage", 
       x = "Year",
       fill = "") +
  theme_minimal()
```