---
title: "Facility visitation exploration"
output: html_document
date: "2023-11-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# load data
```{r, include=FALSE}
facs <- read_csv("facilities.csv")[-1]
missiles <- read_csv("missiles.csv")[-1]
fac_caps_orig <- read_csv("fac_captures.csv")[-1]
miss_caps_orig <- read_csv("miss_captures.csv")[-1]
```

## Per-Facility Visitation Rate
Generate a table with the counts of visits for each facility; visitation leaderboard

First, let's think about how to define a "visit."

Within each flyover, there are often multiple photos taken. To avoid this multi-counting, we want to find a way to identify specific flyovers.
```{r}
length(unique(fac_captures$`Acquisition Date`)) # 1414

length(unique(fac_captures$Mission)) # 269


fac_caps <- fac_caps_orig

fac_caps_grouped <- fac_caps %>% 
  group_by(facility_name, `Acquisition Date`, Mission) %>%
  summarise(n_caps = n()) # 5,018
```

Naively, we could just group by facility name and Mission #. But there's a chance that the same Mission will capture some facilities multiple times.
```{r}
# same mission & facility, diff dates?
fac_caps %>% 
  group_by(facility_name, Mission) %>%
  summarise(n_caps = n()) # 3,945 <- over 1,000 fewer than grouping including Date also, implying the same Mission often photographs the same site on multiple different days

diffs_dates <- fac_caps_grouped %>% ungroup() %>%
  group_by(facility_name, Mission) %>%
  summarize(different_values = n_distinct(`Acquisition Date`) > 1,
            `Acquisition Date`) %>%
  filter(different_values)
```
Looks like it's not uncommon for the same mission to visit the same site twice, sometimes the next day and sometimes months later. 

On the other hand, if we group by just facility name and acquisition date, we run the risk of ignoring cases where multiple different flyovers spotted the same facility on a single day.
```{r}
# same dates & facility, diff mission?
fac_caps %>% 
  group_by(facility_name, `Acquisition Date`) %>%
  summarise(n_caps = n()) # 4,947 <-- 71 fewer than just by date, facility, *and* Mission, implying there are some cases where multiple missions photographed the same site on the same day.

# why is this? Let's check out the cases where it happens.
diffs_missions <- fac_caps_grouped %>% ungroup() %>%
  group_by(facility_name, `Acquisition Date`) %>%
  summarize(different_values = n_distinct(Mission) > 1,
            Mission) %>%
  filter(different_values)
```
On closer inspection, there are some cases where multiple Mission #'s spotted the same facility on the same day, but these seem to have been mostly caused by the same satellite having multiple cameras. For instance, KH-9 satellites have three cameras, which each received a different Mission #. The way this works is that only the first four digits of the Mission number are the actual mission number, but many Mission numbers also have the letter "A" or a "-B" where B is the "bucket number", which appears to correspond to the Camera Type.

There are a few exceptions, however. See, for instance, "Krasnoyarsk-45 Electronchemical Plant - Diffusion (Zelenogorsk in Laughter)" on "1965-01-24"-- both Missions 1016-2 and 4015 seem to have photographed it on that day.

To avoid double-counting, I'll make a new Mission # which only includes the first 4 digits of Mission #, and then group by this Mission #, facility name, and acquisition date.
```{r}
fac_caps <- fac_caps %>%
  mutate(`Abbreviated Mission` = substr(Mission,1,4) )

capture_counts <- fac_caps %>% 
  group_by(facility_name, `Acquisition Date`, `Abbreviated Mission`) %>%
  summarise(n_caps = n()) # 4,951 rows -- only 4 more rows than just grouping by name & date

# let's check out the cases
capture_counts %>% ungroup() %>%
  group_by(facility_name, `Acquisition Date`) %>%
  summarize(different_values = n_distinct(`Abbreviated Mission`) > 1,
            `Abbreviated Mission`) %>%
  filter(different_values)
```
It appears that by grouping by date *and* the first 4 digits of Mission number, we avoid counting 4 events of double flyover as single capture occurences.

Anyways, now we can check out the capture counts.
```{r}
total_counts <- capture_counts %>%
  ungroup() %>%
  group_by(facility_name) %>%
  summarise(n_caps = sum(n_caps))
  
total_counts %>%
  arrange(-n_caps)
```
First glance, this makes a lot of sense! For instance, targets in Russia receive *a lot* more attention than those elsewhere, with some of the least-watched targets being those in Japan.

